---
title: "GO Enrichment Analysis"
author: "Andrea Nieto-Aliseda Sutton"
date: "07/04/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# library(clusterProfiler)
# library(org.Hs.eg.db)

```


#### Summary  

A GO Enrichment Analysis sets out to identify GO terms that are more related to a subset of genes than expected. The expectation of how frequent a GO term should appear is defined by the __universe__, being a definition of a background model.  

The list of genes we are going to use as query has been obtained by the integration between B-ALL associated DMPs falling at enhancer regions as defined by the ensembl regulatory build and PCHi-C data (for now we are using Naive B).  
The genes therefore are those which are possibly being dysregulated by long-range regulatory elements (enhancers).  


#### Clusterprofiler  


```{r, echo=FALSE}


```





```{r universe as captured genes}
# enrichgo <- enrichGO(xg, OrgDb = org.Hs.eg.db, keyType = "ENSEMBL")
# results <- dplyr::as_tibble(as.data.frame(enrichgo@result))
# results <- results[results$p.adjust < 0.05,] #259 
# View(results)
```


```{r universe as all HS genes, echo=FALSE}

```
