---
title: "GO Enrichment Analysis"
author: "Andrea Nieto-Aliseda Sutton"
date: "07/04/2020"
output:
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(clusterProfiler)
library(org.Hs.eg.db)
library(DOSE)
library(tidyverse)
library(vroom)
library(msigdbr)
library(enrichplot)

```


### __Summary__

A GO Enrichment Analysis sets out to identify GO terms that are more related to a subset of genes than expected. The expectation of how frequent a GO term should appear is defined by the __universe__, being a sort of a background model.  

The list of genes we are going to use as query has been obtained by the integration between B-ALL associated DMPs falling at enhancer regions as defined by the ensembl regulatory build and PCHi-C data (for now we are using Naive B).  
The genes therefore are those which are possibly being dysregulated by long-range regulatory elements (enhancers).  
<br>  
 
### __Clusterprofiler__

To start, the subset of genes has to be read in. 
<br>  
```{r, echo=FALSE}
g <- read.table("/media/andrea/SAMSUNG/IJC/Aim3/WORKFLOW-31-03/genes_all.txt", stringsAsFactors = F)
g <- g$V1
```
The subset of genes we want to study consist of __```r length(g)``` genes__
<br>  
```{r, echo=FALSE}

capt <- read.table("/media/andrea/SAMSUNG/IJC/superquery/realfinal_superquery.txt", header = T, stringsAsFactors = F)
baits <- unique(capt$ensembl_gene_id[!is.na(capt$ensembl_gene_id)]) 

```
__```r length(baits)```__ genes will be used as background  
<br>  
```{r, echo=FALSE}

bp <- enrichGO(g, OrgDb = org.Hs.eg.db, keyType = "ENSEMBL", ont = "BP", pAdjustMethod = "fdr", universe = baits)
results1 <- dplyr::as_tibble(as.data.frame(bp@result))

```
The only two GO terms which pass the FDR cutoff < __0.05__ for BP ontology are the following:
```{r, echo=FALSE}

results1[1:2,c(1:3,5:7)]
#write.table(results1, "/media/andrea/SAMSUNG/IJC/Aim3/WORKFLOW-31-03/GO_BP_results.txt", row.names = F, col.names = T, quote = F, sep = "\t")

```
<br>
With a p-value cutoff of __0.1__ the following 3 enriched terms are found:
```{r, echo=FALSE}

bp2 <- enrichGO(g, OrgDb = org.Hs.eg.db, keyType = "ENSEMBL", ont = "BP", pAdjustMethod = "fdr", universe = baits, pvalueCutoff = 0.1)
results2 <- dplyr::as_tibble(as.data.frame(bp2@result))
results2[1:3,c(1:3,5:7)]

```
<br>
With a p-value and q-value cutoff of __0.25__ the following enriched terms are found; 0.25 has been decided as a cutoff given the explanation given here: https://software.broadinstitute.org/cancer/software/gsea/wiki/index.php/FAQ#Why_does_GSEA_use_a_false_discovery_rate_.28FDR.29_of_0.25_rather_than_the_more_classic_0.05.3F
```{r, echo=FALSE}

bp3 <- enrichGO(g, OrgDb = org.Hs.eg.db, keyType = "ENSEMBL", ont = "BP", pAdjustMethod = "fdr", universe = baits, pvalueCutoff = 0.25, qvalueCutoff = 0.25)
results3 <- dplyr::as_tibble(as.data.frame(bp3@result))
results3[1:6,c(1:3,5:7)]

```
<br>  
In essence, the following table shows the GO terms enriched for different cutoff values:  

ID                  | Description                 | p-adjust    | qvalue
--------------------|-----------------------------|-----------------|-----------
GO:0002483      | antigen processing and presentation of endogenous peptide antigen | 0.002837516 | 0.002798439
GO:0019885      | antigen processing and presentation of endogenous peptide antigen via MHC class I | 0.002837516 | 0.002798439
--------------------|-----------------------------|-----------------|-----------
GO:0019883      |	antigen processing and presentation of endogenous antigen                       | 0.084738535 | 0.083571577
--------------------|-----------------------------|-----------------|-----------
GO:0021955      | central nervous system neuron axonogenesis                                      | 0.134774977 | 0.132918952	
GO:0002474      | antigen processing and presentation of peptide antigen via MHC class I          | 0.190618215 | 0.187993157	
GO:0042590	    | antigen processing and presentation of exogenous peptide antigen via MHC class I | 0.246106157 | 0.242716959	

link to MHC wiki: https://en.wikipedia.org/wiki/Major_histocompatibility_complex  
   
<br>
Lets try with a different ontology like Molecular Function (MF) (pvalue < 0.5, qvalue < 0.5)
```{r, echo=FALSE}

mf <- enrichGO(g, OrgDb = org.Hs.eg.db, keyType = "ENSEMBL", ont = "MF", pAdjustMethod = "fdr", universe = baits, pvalueCutoff = 0.5, qvalueCutoff = 0.5)
results4 <- dplyr::as_tibble(as.data.frame(mf@result))
results4[1:10,c(1:3,5:7)]

```
These terms have very low statistical significance  

<br>
Lets try with a different ontology like Cellular Component (CC) (pvalue < 0.5, qvalue < 0.5)
```{r, echo=FALSE}

cc <- enrichGO(g, OrgDb = org.Hs.eg.db, keyType = "ENSEMBL", ont = "CC", pAdjustMethod = "fdr", universe = baits, pvalueCutoff = 0.5, qvalueCutoff = 0.5)
results5 <- dplyr::as_tibble(as.data.frame(cc@result))
results5[1:4,c(1:3,5:7)]

```

<br>  

**What about using transcripts instead of genes to gain statistical significance?**  

```{r echo=FALSE}

t <- read.table("/media/andrea/SAMSUNG/IJC/Aim3/WORKFLOW-31-03/all_transcripts.txt", stringsAsFactors = F)
t <- t$V1
t <- unique(t)

trans <- unique(capt$ensembl_transcript_id[!is.na(capt$ensembl_transcript_id)]) 


bpt <- enrichGO(t, OrgDb = org.Hs.eg.db, keyType = "ENSEMBLTRANS", ont = "BP", pAdjustMethod = "fdr", universe = trans)
results10 <- dplyr::as_tibble(as.data.frame(bpt@result))

#write.table(results10, "/media/andrea/SAMSUNG/IJC/Aim3/WORKFLOW-31-03/GO_BP_trans_results.txt", row.names = F, col.names = T, quote = F, sep = "\t")
```


```{r}
library(GOplot)


r <- results10[,c(1,2,8,6,7,9)]
colnames(r) <- c("ID","Term","Genes","adj_pval","zscore","count")

GOBar(data = r[1:150,])


goplot(bpt, showCategory = 2)
```
<br>
<br>


__What about the Network of Cancer Gene?__  
Network of Cancer Gene (NCG)(A. et al. 2016) is a manually curated repository of cancer genes. NCG release 5.0 (Aug. 2015) collects 1,571 cancer genes from 175 published studies
```{r}
bitr <- bitr(g, "ENSEMBL", "ENTREZID", OrgDb = org.Hs.eg.db)
ncg <- enrichNCG(bitr$ENTREZID, pvalueCutoff = 0.3, minGSSize = 100)
ncg@result[,c(4:7)]
```
Here, count of genes related to LEUKEMIA is the highest (14/78)  

<br>  
<br>  
<br>  
<br> 
















